### 2018-04-17 whitepaper.tex
remove 11/12 usages of the passive voice

\documentclass{bioinfo}
\copyrightyear{2015} \pubyear{2015}

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Manuscript Category}

\begin{document}
\firstpage{1}

\subtitle{Phylogenetics}

\title[Nextstrain]{Nextstrain: real-time tracking of pathogen evolution}
\author[Hadfield \textit{et~al}.]{
James Hadfield\,$^{\text{\sfb 1,}\ast}$,
Colin Megill\,$^{\text{\sfb 1}}$,
Sidney M.~Bell\,$^{\text{\sfb 1,2}}$,
John Huddleston\,$^{\text{\sfb 1,2}}$,
Barney Potter\,$^{\text{\sfb 1}}$,
Charlton Callender\,$^{\text{\sfb 1}}$,
Pavel Sagulenko\,$^{\text{\sfb 3}}$,
Trevor Bedford\,$^{\text{\sfb 1,}\S}$ and
Richard A.~Neher\,$^{\text{\sfb 3,4,5}\S}$
}
\address{
$^{\text{\sf 1}}$Vaccine and Infectious Disease Division, Fred Hutchinson Cancer Research Center, Seattle, WA, USA,
$^{\text{\sf 2}}$Molecular and Cellular Biology Program, University of Washington, Seattle, WA, USA and
$^{\text{\sf 3}}$Max Planck Institute for Developmental Biology, T\"ubingen, Germany.
$^{\text{\sf 4}}$Biozentrum, University of Basel, Basel, Switzerland.
$^{\text{\sf 5}}$SIB Swiss Institute of Bioinformatics, Basel, Switzerland.
}

\corresp{$^\ast$To whom correspondence should be addressed. $^\S$These authors contributed equally to this work.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\abstract{
\textbf{Summary:}
Understanding the spread and evolution of pathogens is important for effective public health measures and surveillance.
Nextstrain consists of a database of viral genomes, a bioinformatics pipeline for phylodynamics analysis, and an interactive visualisation platform.
Together these present a real-time view into the evolution and spread of a range of viral pathogens of high public health importance.
The visualization integrates sequence data with other data types such as geographic information, serology, or host species.
Nextstrain compiles our current understanding into a single accessible location, publicly available for use by health professionals, epidemiologists, virologists and the public alike.
\\
\textbf{Availability and implementation:}
All code (predominantly JavaScript and Python) is freely available from \href{https://github.com/nextstrain}{github.com/nextstrain} and the web-application is available at \href{http://nextstrain.org}{nextstrain.org}.
\\
\textbf{Contact:}
\href{jhadfiel@fredhutch.org}{jhadfiel@fredhutch.org}, \href{tbedford@fredhutch.org}{tbedford@fredhutch.org}, \href{richard.neher@unibas.ch}{richard.neher@unibas.ch}
\\
}

\maketitle

Viral pathogens pose an ever-present danger to global human health, with recent events such as the West-African Ebola epidemic and the ongoing Zika epidemic in the Americas highlighting the risk.
The rapid evolution of these viruses allows inference of epidemic history from genomic data.
Such analyses are often done in isolation, and may lack the spatial or temporal context in which to best interpret the results, which impedes understanding \citep{pybus2013evolutionary}.
Furthermore, the results of analyses are often not made available to the public or health bodies until after publication, which may be too late to effect change in policy or understanding.
We have developed Nextstrain to visualise outbreaks in as close to real time as possible.
Whilst currently encompassing a selection of viruses, extension to non-viral pathogens is forthcoming.

The regularly updated nature and rapidity of these analyses is crucial to the monitoring and understanding of pathogen epidemiology and evolution.
Sequencing times and costs are continually dropping, with on-the-ground sequencing used during the recent West African Ebola and Zika epidemics \citep{quick2016real,faria2017epidemic}.
Rapid methods by which to analyse, interpret, and disseminate results must complement this sepeed of sequencing.

Nextstrain consists of three components:
\textit{Fauna} is a collection of Python scripts to maintain a database representative of all available sequences and related metadata, sourced from public repositories such as NCBI (www.ncbi.nlm.nih.gov), GISAID (www.gisaid.org) and ViPR (www.viprbrc.org), as well as GitHub repositories and other sources of genomic data.
\textit{Augur} is a suite of tools to perform phylodynamic analysis \citep{voltz2013}, including subsampling, alignment, phylogenetic inference, temporal dating of ancestral nodes and discrete trait geographic reconstruction, including inference of the most likely transmission events.
This leverages the maximum likelihood analyses implemented in TreeTime \citep{sagulenko2017treetime}, allowing a full analysis of the entire Ebola epidemic ($n$=1581 genomes) in under 2 hours on a modern laptop.
These scripts separate generic core functionality from a light pathogen-specific layer such that they are easily co-opted to different pathogens.
Visualisation of these analysis is available through \href{http://nextstrain.org}{nextstrain.org}.
This approach is similar in concept to Nextflu \citep{neher2015nextflu} however extended and generalised to different viral pathogens.
There is a growing need for surveillance of non-influenza viruses \citep{tang2017global}, and Nextstrain is able to be extended to most outbreaks with readily accessible genomic data, although we note the potential for recombination or slow mutation rate to confound phylogenetic signal.

\begin{figure}[!tpb]
\centerline{\includegraphics[width=0.48\textwidth]{figure}}
\caption{Genomic epidemiology of Zika virus as of Oct 2017 (live display at \href{http://nextstrain.org/zika}{nextstrain.org/zika}). The main interface consists of three linked panels --- a phylogenetic tree, geographic transmissions, and the genetic diversity across the genome.}
\label{nextstrain}
\end{figure}

\subsection*{Joint temporal and spatial visualisation}
Conveying understanding of pathogen evolution through space and time involves filtering large amounts of data into forms that can be easily reasoned with.
Multiple views into different facets of the data are presented and remain in sync as one interacts with the data.
This allows simultaneous interrogation of phylogenetic and spatial relationships, with additional data such as genotype or serotype expressed through colourings (Figure \ref{nextstrain}).
This is coupled with an interactive time slider to see how the pathogen has evolved and spread over the course of the epidemic.
By animating the temporal dimension, a high level overview of how the entire outbreak unfolded is quickly gained.
This approach both communicates the geographical spread of the epidemic alongside the underlying genomic data that supports this geographic reconstruction.

Maximum likelihood ancestral state reconstruction of discrete traits such as country or region of isolation allows identification of probable transmission events given the sampled data, together with a marginal likelihood for each event.
Internal node colours indicate inferred transmissions on the tree, which are also drawn as links between demes on the map.
It is important to make known the uncertainty of such reconstructions.
The confidence is conveyed by matching colour saturation to the confidence of that trait, and by displaying all relevant information when one hovers over the corresponding branch or isolate on the tree.
Despite this, sampling bias can obscure transmission links, and in cases of extreme bias we have chosen not to display the inferred states.

\subsection*{Monitoring of evolution and adaptation}
Nextstrain tracks and reconstructs mutations across the tree and displays this information as a bar-chart of entropy at each position in the genome, as well as showing the mutations inferred to occur on each branch by hovering over the tree.
Selecting a position in the genome with non-zero entropy reveals the distribution of the segregating variant in the phylogeny and on the map.
This allows interrogation of genetic change which may be adaptive or underlying a change in disease dynamics.

For many pathogens, the emergence and spread of gain-of-function variants is a grave concern.
For instance, China has experienced seasonal epidemics of influenza A/H7N9 over the past five years.
Despite no known cases of human to human transmission, the high mortality rate of $~30\%$ \citep{li2014epidemiology} makes the threat of mutations which facilitate human to human transmission of extreme concern (e.g. see \citep{devries2017three}).
Nextstrain allows monitoring of experimentally determined mutations, informing appropriate interventions and control measures.

\subsection*{A model for public sharing of data}
Nextstrain currently presents a single, continuously updated overview of both endemic viral disease (seasonal influenza, dengue) as well as emergent viral outbreaks (avian influenza, Zika, Ebola), all based upon the same underlying bioinformatics architecture.
This architecture is well positioned to respond to future outbreaks, be they viral or bacterial.

Analysis of such outbreaks relies on public sharing of data, and Nextstrain has the ability to automatically update as new sequences from a range of public databases and repositories appear.
Scientists are justifiably hesitant to cede control of their data, and we try to address these concerns by preventing access to the raw genome sequences, and by clearly indicating the source of each sequence.
Derived data, such as phylogenetic trees, metadata and screenshots are available, and private metadata added via CSV files.
We believe this strikes a compromise between keeping certain data private and allowing the dissemination of results important to the wider scientific community, thereby encouraging collaboration between scientists.
Genomic epidemiology has the potential to inform the public, health organisations and scientists alike, a potential realised by sharing of data in real-time rather than retrospectively \citep{croucher2015application}.

\vspace*{-15pt}

% please add acknowledgment if applicable.
%\section*{Acknowledgements}
%\vspace*{-12pt}

\section*{Funding}

This work was supported by the Open Science Prize to TB and RAN, by the NSF through DGE-1256082 to SMB, by the ERC through StG-260686 to RAN and by NIH R35 GM119774-01 to TB.
TB is a Pew Biomedical Scholar.

\vspace*{-15pt}

\bibliographystyle{bioinfo}
\bibliography{whitepaper}

\end{document}